Ornaments as indicators of Social change before and after European contact at Kiwulan, Northeastern Taiwan

### 2019-06-14 analysis/paper/paper.Rmd
updates based on Ben's suggestions

Q: what is the scale of those exchange network: 
- pre-European contact: middle scale, regional trade network 
- European contact: global scale, international trade network 
- Chinsese contact: small scale, mutual trade network

Q: Is the difference in the frequency of ornaments related to the intensity of contact, does it show any spatial pattern at Kiwulan that indicate the appearance of social inequality.
- pre-European contact: indirect 
- European contact: indirect, colonial context
- Chinsese contact: direct, Han people migration

@@

```{r plot-count-type-by-period}
# tidy the data with categories and shapes
ornaments_shape_count <-
  ornaments_period_clean %>% 
  group_by(period, Categories, Type) %>% 
  tally() %>% 
  filter(!Categories %in% c("Unknown Metal", 
                            "Perforated plate", 
                            "Knited item",
                            "Hair pin",
                            "Bone Ornament",
                            "Metal Piece",
                            "Glass ring",
                            "Golden bead"))

# plot for different shape of agate bead by period
agate_bead_shape <-
  ornaments_period %>% 
  filter(Categories == 'Agate bead') %>%
  ungroup() %>% 
  mutate(Type = ifelse(Type == 'prism/octagonal', 
                       str_remove(Type, '/octagonal'), Type)) %>% 
  mutate(Type = ifelse(Type == 'prism/pentagonal', 
                       str_remove(Type, '/pentagonal'), Type)) %>% 
  mutate(Type = ifelse(Type == 'prism/hexagonal', 
                       str_remove(Type, 'prism/'), Type)) %>%
  mutate(Type = fct_infreq(Type)) %>%
  mutate(Type = fct_rev(Type)) %>% 
  group_by(period, Type) %>% 
  count()

ggplot(agate_bead_shape) +
  geom_col(aes(Type, n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "Frequency",
       x = "Shape of Agate bead") +
  theme_minimal(base_size = 14)

# plot for different shape of metal bell by period
bell_shape <-
  ornaments_period %>% 
  filter(Categories == 'Bell',
         Type != 'unclassified') %>% 
  ungroup() %>% 
  mutate(Type = fct_rev(Type)) %>% 
  group_by(period, Type) %>% 
  count()

ggplot(bell_shape) +
  geom_col(aes(Type, n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "Frequency",
       x = "Shape of Bell") +
  theme_minimal(base_size = 14)

# plot for different shape of metal ring by period
ring_shape <-
  ornaments_period %>% 
  filter(Categories == 'Metal ring') %>%
  ungroup() %>% 
  mutate(Type = fct_infreq(Type)) %>%
  mutate(Type = fct_rev(Type)) %>% 
  group_by(period, Type) %>% 
  count()

ggplot(ring_shape) +
  geom_col(aes(Type, n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "Frequency",
       x = "Shape of Metal ring") +
  theme_minimal(base_size = 14)
```

@@

>>>>

@@

```{r tidy-shape-by-period}
# tidy the data with categories and shapes
ornaments_shape_count <-
  ornaments_period_clean %>% 
  group_by(period, Categories, Type) %>% 
  tally() %>% 
  filter(!Categories %in% c("Unknown Metal", 
                            "Perforated plate", 
                            "Knited item",
                            "Hair pin",
                            "Bone Ornament",
                            "Metal Piece",
                            "Glass ring",
                            "Golden bead"))  %>% 
  ungroup() %>% 
  mutate(Type = as.factor(Type)) 
```

```{r agate-bead-fre-plot}
# plot for different shape of agate bead by period
agate_bead_shape <-
  ornaments_shape_count %>% 
  filter(Categories == 'Agate bead')

agate_bead_shape_plot <- 
ggplot(agate_bead_shape) +
  geom_col(aes(reorder(Type, n), n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "Frequency",
       x = "Shape of Agate bead") +
  theme_minimal(base_size = 14)
```

```{r metal-fre-plot}
# plot for different shape of metal ring by period
ring_shape <-
  ornaments_shape_count %>% 
  filter(Categories == 'Metal ring')

ring_shape_plot <-
ggplot(ring_shape) +
  geom_col(aes(reorder(Type, n), n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "Frequency",
       x = "Shape of Metal ring") +
  theme_minimal(base_size = 14)
```

```{r bell-fre-plot}
# plot for different shape of metal bell by period
bell_shape <-
  ornaments_shape_count %>% 
  filter(Categories == 'Bell',
         Type != 'unclassified')

bell_shape_plot <-
ggplot(bell_shape) +
  geom_col(aes(reorder(Type,n), n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "Frequency",
       x = "Shape of Bell") +
  theme_minimal(base_size = 14)
```

```{r glass-bead-fre-plot}
# plot for different shape of glass bead by period
glass_bead_shape <-
  ornaments_shape_count %>% 
  filter(Categories == 'Glass bead')

glass_bead_plot <-
ggplot(glass_bead_shape) +
  geom_col(aes(Type, n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "Frequency",
       x = "Shape of Glass Bead") +
  theme_minimal(base_size = 14)
```

```{r}
library(cowplot)
plot_grid(agate_bead_shape_plot, ring_shape_plot,
          labels = c("A", "B"))

plot_grid(bell_shape_plot, glass_bead_plot,
          labels = c("C", "D"))
```

@@

Q: what is the scale of those exchange network: 
- pre-European contact: middle scale, regional trade network 
- European contact: global scale, international trade network 
- Chinese contact: small scale, mutual trade network

Q: Is the difference in the frequency of ornaments related to the intensity of contact, does it show any spatial pattern at Kiwulan that indicate the appearance of social inequality.
- pre-European contact: indirect 
- European contact: indirect, colonial context
- Chinese contact: direct, Han people migration
